Update of the ozone dry deposition in the OsloCTM3 v1.0

### 2018-11-21 DryDep.tex
Found some mistake in the analysis. Re-did the plots, but haven't had the nerve to go through the text again...

\begin{table*}[t]
  \caption{Summary of specifications of all simulations discussed in this section. For simplicity, only the tested parameters are listed. An x denotes that the model was run exactly in the configuration as has been described in Section~\ref{sec:model_des}.}
  \begin{tabular}{lccccccccc}
    \tophline
    \multirow{2}{*}{Simulation} & Wesely scheme & \multicolumn{3}{c}{EMEP scheme} & \multicolumn{2}{c}{Greening season} & $v^\chem{O_3}_\text{ice/snow}$ & Forcings\\
    & & $f_\text{phen}$ & $f_\text{light}$ & $f_\text{SW}$ & SMOKE-BEIS & $5\,\unit{^\circ C}$-days & (\unit{m\,s^{-1}}) & (year)\\
    \middlehline
    Wesely\_type      & x & \multicolumn{3}{c}{n/a}  & x & n/a   & 1/2000  & 2014 \\
    EMEP\_full        &   & incl. TR & x & SWLV1     & x &       & 1/2000  & 2014 \\
    EMEP\_offLight    &   & incl. TR & 1 & SWLV1     & x &       & 1/2000  & 2014 \\
    EMEP\_offPhen     &   & 1        & x & SWLV1     & x &       & 1/2000  & 2014 \\
    EMEP\_SWVL4       &   & incl. TR & x & SWLV4     & x &       & 1/2000  & 2014 \\
    EMEP\_ppgs        &   & incl. TR & x & SWLV1     &   & NH    & 1/2000  & 2014 \\
    EMEP\_ppgs\_2005  &   & incl. TR & x & SWLV1     &   & NH    & 1/2000  & 2005 \\
    EMEP\_ppgssh      &   & excl. TR & x & SWLV1     &   & NH+SH & 1/2000  & 2014 \\
    EMEP\_ppgssh\_ice &   & excl. TR & x & SWLV1     &   & NH+SH & 1/10000 & 2014 \\
    \bottomhline
  \end{tabular}
  \label{tab:simsum}
  %\belowtable{} % Table Footnotes
\end{table*}
%
\begin{figure}[t]
  \includegraphics[width=8.3cm]{pictures/final-ozone_diff_maps}
  \caption{Relative difference between the sensitivity simulation \emph{EMEP\_ppgsssh} and the \emph{Wesely\_type} in (a) total surface ozone; (b) average ozone dry deposition velocity; (c) total amount of ozone removed from the atmosphere by dry deposition.}
  \label{fig:diff_maps}
\end{figure}
%
In Fig.~\ref{fig:diff_maps}, we show the average relative difference between \emph{EMEP\_ppgsssh} and \emph{Wesely\_type} on global maps by means of surface ozone burden, dry deposition velocity and total ozone dry deposition. Surface ozone increases globally except for regions covered by tropical forest. Especially in desert regions in Africa and Asia, surface ozone increases by up to $100\,\unit{\%}$. Consistently, dry deposition velocities decrease globally (up to $100\,\unit{\%}$) in desert regions, while they increase over tropical forest. With respect to total dry deposition, the picture is a bit less clear. We find a decrease of total dry deposition of ozone in desert regions and ocean covered areas and an increase in regions covered by tropical forest, while at mid and high latitudes in both hemispheres only small changes are visible. A possible explanation to these changes especially in desert regions is the difference between the previously prescribed dry deposition velocities in the Wesely scheme and the adjusting dry deposition velocities with respect to meteorology in the EMEP scheme. We cover this in more detail in the following section.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Comparison with multi-model evaluation}
\label{subsec:model}
%
In our evaluation of results, we closly follow suggestions by \citet{ACP:Hardacre2015}. For the purpose of comparison with the multi-model mean of the therein participating Task Force on Hemispheric Transport of Air Pollution (TF HTAP) models, we also have re-grided our data to a horizontal resolution of $3^\circ\times 3^\circ$. In Section~\ref{subsubsec:zonal}, we look at zonal ditributions of $[\chem{O_3}](z_0)$, $v^\chem{O_3}_\text{DD}$, and $\sum\chem{O_3^\text{DD}}$ for all our senstivity simulations and study seasonal cycles of hemispheric ozone as well as nine land surface types which have been retrospectively seperated (Section~\ref{subsubsec:seasons}). From this, we estimate the total annual ozone dry deposition onto ocean, ice, and land surfaces and compare with results from \citet{ACP:Luhar2017}.
%
\subsubsection{Zonal distribution}
\label{subsubsec:zonal}
%
The annual zonal average with respect to surface ozone concentration (Fig.~\ref{fig:mmm_drydep}a) displays, in consistancy with Fig.~\ref{fig:diff_maps}a), a general, global increase of ozone by up to $10\,\unit{ppb}$ which is largest between $(25-50)\,\unit{^\circ N}$ including the major deserts. In the deep tropics ($5\,\unit{^\circ S}-5\,\unit{^\circ N}$), the increase is smallest ($\Delta[\chem{O_3}]\leq 5\,\unit{ppb}$). We find that the EMEP scheme further intensifies the strong asymmetry between northern and southern hemisphere as a consequence of the distribution of the continental land masses and vegetation thereon. Among the sensitivity studies, there is only a low variance. Most remarkable and as expected due to the strong decrease in prescribed dry deposition velocity over ice and snow, \emph{EMEP\_ppgssh\_ice} displays almost a doubling of surface ozone in the high Arctics compared to \emph{Wesely\_type} but affects ozone concentrations already at about $50\,\unit{^\circ}$ in both hemispheres. Neglecting any dependence on solar radiation in the stomatal conductance formulation (\emph{EMEP\_offLight}) -- or in other words allowing photosynthesis 24/7 -- generally decreases the ozone concentration by about $2\,\unit{ppb}$, while choosing soil water at deeper levels than the surface layer (\emph{EMEP\_SWVL4}) increases $[\chem{O_3}](z_0)$ slightly. Rather surprisingly, switching of phenology completely (\emph{EMEP\_offPhen}), doesn't make much difference. The scenario of differing emissions (2005 in comparison to 2014 or more specifically \emph{EMEP\_ppgs\_2005} compared to \emph{EMEP\_ppgs}), yields higher ozone concentrations in the northern hemisphere in 2005 in accordance to a reduction in sulfur and \chem{NO_x} emissions in south east Asia in later years. An opposite tendency is seen for latitudes south of $30\,\unit{^\circ N}$, propbably owing to the industrial development of countries in these regions.

In Fig.~\ref{fig:mmm_drydep}b), we show the corresponding average dry deposition velocities, as reconstructed from the ratio between surface concentration and total dry deposition since we have no direct output of these. The $v^\chem{O_3}_\text{DD}$ in the EMEP scheme are much smaller then in the Wesely scheme and about a factor of $3$ smaller compared to the multi-model-mean for latitudes between $50\,\unit{^\circ S}$ and $70\,\unit{^\circ N}$. In the remaining latitude bands, except for \emph{EMEP\_ppgssh\_ice}, all realisations of the EMEP scheme are constantly above the multi-model-mean. This indicates, that with respect to other models, the updated dry deposition velocity above ice and snow should be considered as new standard for the OsloCTM3. The dry deposition velocities are of cause independent of the emission scenario, but display a strong dependence on $f_\text{light}$ and $f_\text{phen}$. The closer to $1$ these factores are the higher do the dry deposition velocities become (compare \emph{EMEP\_offLight}, \emph{EMEP\_offPhen}, and \emph{EMEP\_ppgssh}). That highlights the importance of a proper PAR from meteorological input in CTMs or radiative transfer in PAR in global climate models (GCM) for canopy conductance computation. Albeit, the total numbers are smaller, the shape of the zonal average dry deposition velocities from the EMEP scheme and the multi-model-mean are similar (see figure in Supplement~S.6.1). The biggest exception is the zonal band $(12-30)\,\unit{^\circ S}$, which coincide with the location of Australia and its large desert regions.

The annual total ozone dry deposition is shown in Fig.~\ref{fig:mmm_drydep}b). In accordance to the priviously decribed features, we observe a reduction of the global total ozone dry deposition to almost half of \emph{Wesely\_type} in any of the sensitivity studies. The occurence of this reduction in the zonal bands where the major deserts are located may point to a important change in $v_\text{desert}^\chem{O_3}$. Consulting the old parameter file, we find $v_\text{desert}^\chem{O_3} \equiv v_\text{tundra}^\chem{O_3} = 0.26\,\unit{cm\,s^{-1}}$, while in the EMEP scheme $v_\text{desert}^\chem{O_3}=0.05\,\unit{cm\,s^{-1}}$. Similarly, dry deposition velocities over ice and snow and ocean have been even higher in the Wesely scheme than in the original EMEP scheme ($v_\text{ice/snow}^\chem{O_3}\equiv v_\text{water}^\chem{O_3}=0.07\,\unit{cm\,s^{-1}}$). These differences in baseline dry deposition velocities over huge parts of the unvegetated surface of the Earth (see Fig.~\ref{fig:pft_landsurface}) amount to most of the quantitative change we see between the Wesely and the EMEP scheme. We further elaborate on this in the following section.

\begin{figure}[t]
  \includegraphics[width=8.3cm]{pictures/final-vmr_surfaceozone}\\
  \includegraphics[width=8.3cm]{pictures/final-total_ozone_drydepvelo}\\
  \includegraphics[width=8.3cm]{pictures/final-total_ozone_drydep}
  \caption{Comparison of the manifold of OsloCTM3 integrations with respect to (a) Surface ozone concentrations, (b) Annual average ozone dry deposition velocity, (c) Total annual ozone dry deposition. The different colors indicate sets of simulation with similar baselines. The multi-model mean from the evaluation of HTAP models by \citet{ACP:Hardacre2015} is shown as a reference (where available).}
  \label{fig:mmm_drydep}
\end{figure}
%
In Fig.~\ref{fig:mmm_drydep_season}a), the average zonal ozone dry deposition is shown seperated by month. Where available, we have added the multi-model-mean given by \citet{ACP:Hardacre2015} as reference. There are two major phases apparent, NH and SH greening season. Former is reflected in a pronounced peak of $v_\text{DD}^\chem{O_3}$ in the northern mid latitudes, while the absence of this peak can be attributed to the latter in addition to a southward shift of the tropical peak dry deposition velocity. In the Wesely scheme, NH midlatitude peak velocities appear in June compared to July in the EMEP scheme, indicating that the prescribed seasonal cycles differ. The corresponding total monthly ozone dry deposition is shown in Fig.~\ref{fig:mmm_drydep_season}b). In general, the seasonal patters are quite similar in the Wesely scheme and the EMEP scheme, displaying a strong symmetry around $10\,\unit{^\circ N}$ in January/February and November/December, respectively. What differs most is the molding and intensitity of the NH peak dry deposition.Both schemes reach the maximum in June/July but the peak is much more differenciated in March already in the Wesely scheme. Similarly, the SH tropical peak dry deposition is reached in August/September but sustained longer into October in the Wesely scheme. Since we have not conducted any simulation with a meteorological year other than 2005, we cannot elaborate on whether this is a special feature of our chosen year or not.
%
\begin{figure}[t]
  \includegraphics[width=8.3cm]{pictures/final-total_ozone_drydepvelo_monthly}\\
  \includegraphics[width=8.3cm]{pictures/final-total_ozone_drydep_monthly}
  \caption{Comparison of the manifold of OsloCTM3 integrations with respect to (a) Zonal average ozone dry deposition velocities; (b) Total annual amount of ozone removed from the atmosphere via dry deposition. The multi-model mean from the evaluation of HTAP models by \citet{ACP:Hardacre2015} is shown as a reference (where available).}
  \label{fig:mmm_drydep_season}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Average seasonal cycles}
\label{subsubsec:seasons}
%
To disentangle the contributions of different regions to the global ozone budget further, we will briefly look at different projections of seasonal cycles.

In Fig.~\ref{fig:mmm_drydep_hem}, the total annual ozone dry deposition seperated into mid and high latitudes in the northern hemisphere ($(30-90)\,\unit{^\circ N}$), the tropics and subtropics ($30\,\unit{^\circ S}-30\,\unit{^\circ N}$), and the mid and high latitudes in the southern hemisphere ($(30-90)\,\unit{^\circ S}$) are shown. As reference, we have again added the multi-model-mean by \citet{ACP:Hardacre2015}. As expected, the NH mid and high latitudes display a strongly pronounced seasonal cycle, while it is less pronounced in the tropics and in the SH. The highest ozone dry deposition is found in the tropics and amounts on average to the  peak levelof dry deposition in the NH for the multi-model-mean \citep{ACP:Hardacre2015} and EMEP scheme. The average tropical ozone dry deposition in the Wesely scheme diverges by about $5\,\unit{Tg}$ in comparison to its NH maximum. The seasonal cycle in the EMEP scheme appears to be shifted towards later in the year in the NH. While the general seasonal cycle in the tropics and subtropics is not changing, we find the strongest sensitivity to the verious realisations of the EMEP scheme there, with \emph{EMEP\_offLight} and \emph{EMEP\_ppgs\_2005} the highest and the lowest, respectively. This indicates that surface ozone in the tropics may be very sensitive to changes to emissions and changes in longwave radiative transfer (e.g., cloud cover, aerosol).
  \includegraphics[width=12cm]{pictures/final-total_ozone_drydep_hem}
  \caption{Seasonal cycle of total annual amount of ozone removed from the atmosphere through dry deposition separated into northern hemisphere (NH), tropics (TR), and southern hemisphere (SH). The multi-model mean from the evaluation of HTAP models by \citet{ACP:Hardacre2015} is shown as a reference.}
  \label{fig:mmm_drydep_hem}
As suggested by \citet{ACP:Hardacre2015}, we retrospectively separate ozone dry deposition velocities with respect to surface types. Based on CML2 dynamic land surface, we generated masks for 9 different surface types (Fig.~\ref{fig:pft_landsurface}a) and used these to select gridboxes with a high percentage of these surface types, ranging from $22\,\unit{\%}$ only for tundra in the southern hemisphere to $98\,\unit{\%}$ regarding ocean. Since we have not perform a full unfolding on the data, the results should be regarded with suspicion.
  \includegraphics[width=12cm]{pictures/final-total_ozone_drydepvelo_monthly_pft_nh}
  \caption{Average ozone dry deposition velocities separated by land use type and compared to results from \citep{ACP:Hardacre2015}.}
  \label{fig:mmm_drydep_season_pft}
The Wesely scheme uses $v^\chem{O_3}_\text{tundra/desert} = 0.26\,\unit{cm\,s^{-1}}$ \citep{JGR:Hough1991}, while in the EMEP scheme we use a base value of $v^\chem{O_3}_\text{desert} = 0.05\,\unit{cm\,s^{-1}}$ and $v^\chem{O_3}_\text{tundra} = 0.24\,\unit{cm\,s^{-1}}$. From measurements of ozone fluxes in the Sahara desert, $v^\chem{O_3}_\text{desert, day} = 0.1\,\unit{cm\,s^{-1}}$, $v^\chem{O_3}_\text{desert, night} = 0.04\,\unit{cm\,s^{-1}}$, and $\left<v^\chem{O_3}_\text{desert}\right> = 0.065\,\unit{cm\,s^{-1}}$ have been deduced \citep{AE:Gusten1995}. This implies, that ozone dry deposition over desert regions is highly overestimated in the Wesely scheme, while it may be underestimated in the EMEP scheme. Similarly, the dry deposition over water $v^\chem{O_3}_\text{water}=0.07\,\unit{cm\,s^{-1}}$ in the Wesely scheme, while in the EMEP scheme we use a base value of $v^\chem{O_3}_\text{water} = 0.05\,\unit{cm\,s^{-1}}$. $v^\chem{O_3}_\text{water} = 0.019\,\unit{cm\,s^{-1}}$ \citep{JGR:Helmig2012}. \citet{ACP:Luhar2017} found $0.018\,\unit{cm\,s^{-1}} \leq \left<v^\chem{O_3}_\text{water}\right> \leq 0.039\,\unit{cm\,s^{-1}}$.
\begin{table*}[t]
\caption{Total ozone dry deposition for the respective year in \unit{Tg\,a^{-1}} separated into ocean, ice and, land contributions.}
\begin{tabular}{lcccccccccccc}
\multirow{3}{*}{Simulation} & \multicolumn{3}{c}{Ocean} & \multicolumn{3}{c}{Ice} & \multicolumn{3}{c}{Land} & \multicolumn{3}{c}{Total}\\
& NH & SH & Global & NH & SH & Global & NH & SH & Global & NH & SH & Global\\
& \multicolumn{3}{c}{(\unit{Tg\,a^{-1}})} & \multicolumn{3}{c}{(\unit{Tg\,a^{-1}})} & \multicolumn{3}{c}{(\unit{Tg\,a^{-1}})} & \multicolumn{3}{c}{(\unit{Tg\,a^{-1}})}\\
Wesely\_type & 88.7 & 110.3 & 199.0 & 0.7 & 5.0 & 5.7 & 346.1 & 153.5 & 499.6 & 612.4 & 347.9 & 960.2 \\
EMEP\_full & 67.1 & 84.7 & 151.8 & 0.6 & 4.4 & 5.0 & 236.0 & 112.3 & 348.2 & 408.6 & 245.6 & 654.2 \\
EMEP\_offLight & 66.6 & 84.1 & 150.7 & 0.6 & 4.4 & 5.0 & 267.8 & 129.8 & 397.6 & 451.3 & 267.2 & 718.6 \\
EMEP\_offPhen & 67.1 & 84.6 & 151.7 & 0.6 & 4.4 & 5.0 & 239.9 & 116.0 & 355.9 & 413.9 & 249.9 & 663.7 \\
EMEP\_SWVL4 & 68.1 & 86.1 & 154.2 & 0.6 & 4.5 & 5.1 & 241.8 & 115.0 & 356.8 & 417.2 & 250.7 & 667.9 \\
EMEP\_ppgs & 67.1 & 84.6 & 151.7 & 0.6 & 4.4 & 5.0 & 238.7 & 112.6 & 351.3 & 411.4 & 246.0 & 657.4 \\

>>>>

\begin{table*}[t]
  \caption{Summary of specifications of all simulations discussed in this section. For simplicity, only the tested parameters are listed. An x denotes that the model was run exactly in the configuration as has been described in Section~\ref{sec:model_des}.}
  \begin{tabular}{lccccccccc}
    \tophline
    \multirow{2}{*}{Simulation} & Wesely scheme & \multicolumn{3}{c}{EMEP scheme} & \multicolumn{2}{c}{Greening season} & $v^\chem{O_3}_\text{ice/snow}$ & Forcings\\
    & & $f_\text{phen}$ & $f_\text{light}$ & $f_\text{SW}$ & SMOKE-BEIS & $5\,\unit{^\circ C}$-days & (\unit{m\,s^{-1}}) & (year)\\
    \middlehline
    Wesely\_type      & x & \multicolumn{3}{c}{n/a}  & x & n/a   & 1/2000  & 2014 \\
    EMEP\_full        &   & incl. TR & x & SWLV1     & x &       & 1/2000  & 2014 \\
    EMEP\_offLight    &   & incl. TR & 1 & SWLV1     & x &       & 1/2000  & 2014 \\
    EMEP\_offPhen     &   & 1        & x & SWLV1     & x &       & 1/2000  & 2014 \\
    EMEP\_SWVL4       &   & incl. TR & x & SWLV4     & x &       & 1/2000  & 2014 \\
    EMEP\_ppgs        &   & incl. TR & x & SWLV1     &   & NH    & 1/2000  & 2014 \\
    EMEP\_ppgs\_2005  &   & incl. TR & x & SWLV1     &   & NH    & 1/2000  & 2005 \\
    EMEP\_ppgssh      &   & excl. TR & x & SWLV1     &   & NH+SH & 1/2000  & 2014 \\
    EMEP\_ppgssh\_ice &   & excl. TR & x & SWLV1     &   & NH+SH & 1/10000 & 2014 \\
    \bottomhline
  \end{tabular}
  \label{tab:simsum}
  %\belowtable{} % Table Footnotes
\end{table*}
%
\begin{figure}[t]
  \includegraphics[width=8.3cm]{pictures/final-ozone_diff_maps}
  \caption{Relative difference between the sensitivity simulation \emph{EMEP\_ppgsssh} and the \emph{Wesely\_type} in (a) total surface ozone; (b) average ozone dry deposition velocity; (c) total amount of ozone removed from the atmosphere by dry deposition.}
  \label{fig:diff_maps}
\end{figure}
%
In Fig.~\ref{fig:diff_maps}, we show the average relative difference between \emph{EMEP\_ppgsssh} and \emph{Wesely\_type} on global maps by means of surface ozone burden, dry deposition velocity and total ozone dry deposition. Surface ozone increases globally except for regions covered by tropical forest. Especially in desert regions in Africa and Asia, surface ozone increases by up to $100\,\unit{\%}$. Consistently, dry deposition velocities decrease globally by yhe same order of magnitude in these regions, while they increase over tropical forest. With respect to total dry deposition, the picture is a bit less clear. We find a decrease of total dry deposition of ozone in desert regions and ocean covered areas and an increase in regions covered by tropical forest, while at mid and high latitudes in both hemispheres only small changes are visible. A possible explanation to these changes especially in desert regions is the difference between the previously prescribed dry deposition velocities in the Wesely scheme and the adjusting dry deposition velocities with respect to meteorology in the EMEP scheme. We cover this in more detail in the following section.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Comparison with modeling results}
\label{subsec:model}
%
In our evaluation of results, we closly follow suggestions by \citet{ACP:Hardacre2015}. For the purpose of comparison with the multi-model mean of the therein participating Task Force on Hemispheric Transport of Air Pollution (TF HTAP) models, we also have re-grided our data to a horizontal resolution of $3^\circ\times 3^\circ$. In Section~\ref{subsubsec:zonal}, we look at zonal ditributions of $[\chem{O_3}](z_0)$, $v^\chem{O_3}_\text{DD}$, and $\sum\chem{O_3^\text{DD}}$ for all our senstivity simulations and study seasonal cycles of hemispheric ozone as well as nine land surface types which have been retrospectively seperated (Section~\ref{subsubsec:seasons}). From this, we estimate the total annual ozone dry deposition onto ocean, ice, and land surfaces and compare with results from \citet{ACP:Luhar2017}.
%
\subsubsection{Zonal distribution}
\label{subsubsec:zonal}
%
The annual zonal average with respect to surface ozone concentration (Fig.~\ref{fig:mmm_drydep}a) displays, in consistancy with Fig.~\ref{fig:diff_maps}a), a general, global increase of ozone by up to $10\,\unit{ppb}$ which is largest between $(25-50)\,\unit{^\circ N}$ including the major deserts. In the deep tropics ($5\,\unit{^\circ S}-5\,\unit{^\circ N}$), the increase is smallest ($\Delta[\chem{O_3}]\leq 5\,\unit{ppb}$). We find that the EMEP scheme further intensifies the strong asymmetry between northern and southern hemisphere as a consequence of the distribution of the continental land masses and vegetation thereon. Among the sensitivity studies, there is only a low variance. Most remarkable and as expected due to the strong decrease in prescribed dry deposition velocity over ice and snow, \emph{EMEP\_ppgssh\_ice} displays almost a doubling of surface ozone in the high Arctics compared to \emph{Wesely\_type} but affects ozone concentrations down to latitudes at about $50\,\unit{^\circ}$ in both hemispheres. Neglecting any dependence on solar radiation in the stomatal conductance formulation (\emph{EMEP\_offLight}) -- or in other words allowing photosynthesis 24/7 -- generally decreases the ozone concentration by about $2\,\unit{ppb}$, while choosing soil water at deeper levels than the surface layer (\emph{EMEP\_SWVL4}) increases $[\chem{O_3}](z_0)$ slightly. Rather surprisingly, switching off the phenology completely (\emph{EMEP\_offPhen}), doesn't make much difference. The scenario of differing emissions (2005 in comparison to 2014 or more specifically \emph{EMEP\_ppgs\_2005} compared to \emph{EMEP\_ppgs}), yields higher ozone concentrations in the northern hemisphere in 2005 in accordance to a reduction in sulfur and \chem{NO_x} emissions in south east Asia in later years. An opposite tendency is seen for latitudes south of $30\,\unit{^\circ N}$, propbably owing to the industrial development of countries in these regions.

Average dry deposition velocities have been reconstructed from the ratio between surface concentration and total dry deposition
\begin{equation}
  v^\chem{O_3}_\text{DD} = \frac{\sum \chem{O_3^\text{DD}}(z_0)}{\chem{O_3}(z_0)}\cdot c_\text{month},
\end{equation}
with $c_\text{month} = \frac{\text{layer thickness}}{\text{seconds in month}}$. This way of computing dry deposition velocities has certain drawbacks of which one is the freedom of choice of the \emph{layer thickness}. We chose the average thickness of the lowermost model level ($16\cdot10^{2}\,\unit{cm}$) to match the prescribed dry deposition velocities above ice and snow ($v_\text{ice/snow}^\chem{O_3} = 0.05\,\unit{cm\,s^{-1}}$). The resulting $v^\chem{O_3}_\text{DD}$ are shown in Fig.~\ref{fig:mmm_drydep}b). The dry deposition velocities in the EMEP scheme are much smaller than in the Wesely scheme for latitudes between $50\,\unit{^\circ S}$ and $70\,\unit{^\circ N}$. In the Arctics, except for \emph{EMEP\_ppgssh\_ice}, all realisations of the EMEP scheme are constantly above the multi-model-mean. This indicates, that with respect to other models, the updated dry deposition velocity above ice and snow should be considered as new standard for the OsloCTM3. The dry deposition velocities are of cause independent of the emission scenario, but display a strong dependence on $f_\text{light}$ and $f_\text{phen}$. The closer to $1$ these factores are the higher the dry deposition velocities become (compare \emph{EMEP\_offLight}, \emph{EMEP\_offPhen}, and \emph{EMEP\_ppgssh}). That highlights the importance of a proper PAR from meteorological input in CTMs or radiative transfer in PAR in global climate models (GCM) for canopy conductance computation. Albeit, the total numbers are smaller, the shape of the zonal average dry deposition velocities from the EMEP scheme and the multi-model-mean are similar (for surface tupe partitioning see figure in Supplement~S.6.1). The biggest exception is the zonal band $(12-30)\,\unit{^\circ S}$, which coincide with the location of Australia and its large desert regions, and its counterpart in the northern hemisphere.

The annual total ozone dry deposition is shown in Fig.~\ref{fig:mmm_drydep}b). In accordance to the priviously decribed features, we observe a reduction of the global total ozone dry deposition to almost half of \emph{Wesely\_type} in all sensitivity studies. The occurence of this reduction in the zonal bands where the major deserts are located may point to a important change in $v_\text{desert}^\chem{O_3}$. Consulting the Wsely scheme parameter file, we find $v_\text{desert}^\chem{O_3} \equiv v_\text{tundra}^\chem{O_3} = 0.26\,\unit{cm\,s^{-1}}$ \citep{JGR:Hough1991}, while in the EMEP scheme $v_\text{desert}^\chem{O_3}=0.05\,\unit{cm\,s^{-1}}$ and $v^\chem{O_3}_\text{tundra} = 0.24\,\unit{cm\,s^{-1}}$, respectively. Similarly, dry deposition velocities over ice and snow and ocean have been even higher in the Wesely scheme ($v_\text{ice/snow}^\chem{O_3}\equiv v_\text{water}^\chem{O_3}=0.07\,\unit{cm\,s^{-1}}$) than in the original EMEP scheme ($v_\text{ice/snow}^\chem{O_3}\equiv v_\text{water}^\chem{O_3}=0.05\,\unit{cm\,s^{-1}}$). These differences in baseline dry deposition velocities over huge parts of the unvegetated surface of the Earth (see Fig.~\ref{fig:pft_landsurface}) amount to most of the quantitative difference between the Wesely and the EMEP scheme. We further elaborate on this in the following section.

\begin{figure}[t]
  \includegraphics[width=8.3cm]{pictures/final-vmr_surfaceozone}\\
  \includegraphics[width=8.3cm]{pictures/final-total_ozone_drydepvelo}\\
  \includegraphics[width=8.3cm]{pictures/final-total_ozone_drydep}
  \caption{Comparison of the manifold of OsloCTM3 integrations with respect to (a) Surface ozone concentrations, (b) Annual average ozone dry deposition velocity, (c) Total annual ozone dry deposition. The different colors indicate sets of simulation with similar baselines. The multi-model mean from the evaluation of HTAP models by \citet{ACP:Hardacre2015} is shown as a reference (where available).}
  \label{fig:mmm_drydep}
\end{figure}
%
In Fig.~\ref{fig:mmm_drydep_season}a), the average zonal ozone dry deposition is shown seperated by month. Where available, we have added the multi-model-mean given by \citet{ACP:Hardacre2015} as reference. There are two major phases apparent, NH and SH greening season. Former is reflected in a pronounced peak of $v_\text{DD}^\chem{O_3}$ in the northern mid latitudes, while the absence of this peak can be attributed to the latter in addition to a southward shift of the tropical peak dry deposition velocity. In the Wesely scheme, NH midlatitude peak velocities appear in June compared to July in the EMEP scheme, indicating that the prescribed seasonal cycles differ. The corresponding total monthly ozone dry deposition is shown in Fig.~\ref{fig:mmm_drydep_season}b). In general, the seasonal patterns are quite similar in the Wesely scheme and the EMEP scheme, displaying a strong symmetry around $10\,\unit{^\circ N}$ in January/February and November/December, respectively. What differs most is the molding and intensitity of the NH peak dry deposition. Both schemes reach the maximum in June/July but the peak is much more differenciated in March already in the Wesely scheme. Similarly, the SH tropical peak dry deposition is reached in August/September but sustained longer, into October, in the Wesely scheme. Since we have not conducted any simulation with a meteorological year other than 2005, we cannot elaborate on whether this is a special feature of our chosen year or not.
%
\begin{figure}[t]
  \includegraphics[width=8.3cm]{pictures/final-total_ozone_drydepvelo_monthly}\\
  \includegraphics[width=8.3cm]{pictures/final-total_ozone_drydep_monthly}
  \caption{Comparison of the manifold of OsloCTM3 integrations with respect to (a) Zonal average ozone dry deposition velocities; (b) Total annual amount of ozone removed from the atmosphere via dry deposition. The multi-model mean from the evaluation of HTAP models by \citet{ACP:Hardacre2015} is shown as a reference (where available).}
  \label{fig:mmm_drydep_season}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Average seasonal cycles}
\label{subsubsec:seasons}
%
To further disentangle the contributions of different regions to the global ozone budget, we will briefly look at different projections of seasonal cycles.

In Fig.~\ref{fig:mmm_drydep_hem}, the total annual ozone dry deposition seperated into mid and high latitudes in the northern hemisphere ($(30-90)\,\unit{^\circ N}$), the tropics and subtropics ($30\,\unit{^\circ S}-30\,\unit{^\circ N}$), and the mid and high latitudes in the southern hemisphere ($(30-90)\,\unit{^\circ S}$) are shown. As reference, we have added the multi-model-mean by \citet{ACP:Hardacre2015}. As expected, the NH mid and high latitudes display a strongly pronounced seasonal cycle, while it is less pronounced in the tropics (due to the lack of seasons) and in the SH (due to the small percentage of vegetated surface). The highest ozone dry deposition is found in the tropics and amounts on average to the  peak level of dry deposition in the NH for the multi-model-mean \citep{ACP:Hardacre2015} and EMEP scheme. In the Wesely scheme, the average tropical ozone dry deposition diverges by about $5\,\unit{Tg}$ in comparison to its corresponding NH maximum. The seasonal cycle in the EMEP scheme appears to be shifted towards later in the year in the NH. While the general seasonal cycle in the tropics and subtropics is not changing, we find the strongest sensitivity to the various realisations of the EMEP scheme there, with \emph{EMEP\_offLight} and \emph{EMEP\_ppgs\_2005} the highest and the lowest, respectively. This indicates that surface ozone in the tropics may be very sensitive to changes to emissions and changes in longwave radiative transfer (e.g., cloud cover, aerosol).

%
\begin{figure*}[t]
  \includegraphics[width=12cm]{pictures/final-total_ozone_drydep_hem}
  \caption{Seasonal cycle of total annual amount of ozone removed from the atmosphere through dry deposition separated into northern hemisphere (NH), tropics (TR), and southern hemisphere (SH). The multi-model mean from the evaluation of HTAP models by \citet{ACP:Hardacre2015} is shown as a reference.}
  \label{fig:mmm_drydep_hem}
\end{figure*}
%
As suggested by \citet{ACP:Hardacre2015}, we retrospectively separate ozone dry deposition velocities with respect to surface types. Based on a CLM~2 average dynamic land surface map, we generated masks for 9 different surface types (Fig.~\ref{fig:pft_landsurface}a) and used these to select gridboxes with a high percentage of these surface types, ranging from a meager $22\,\unit{\%}$ for tundra in the southern hemisphere to $98\,\unit{\%}$ regarding ocean. From this CLM~2 based map, it was not possible to select gridboxes with $100\,\unit{\%}$ for each surface type. Since we have not perform a full unfolding on the data, the results should be treated with slight caution. In Fig.~\ref{fig:mmm_drydep_season_pft}, the seasonal cycles of dry deposition velocities are shown for the nine surface categories. The patterns and absolute numbers differ between the schemes and the multi-model-mean. Exceptually large is the divergence for desert regions. The multi-model-mean displays a very pronounced seasonal cycle ($0.15\,\unit{cm\,s^{-1}} \leq v^\chem{O_3}_\text{desert}\leq 0.10\,\unit{cm\,s^{-1}}$). The Wesely scheme as well as the EMEP scheme do not indicate any seasonal cycle. The estimated dry deposition velocities over desert regions are consistent with the previously mentioned baseline values of ozone dry deposition,  about one order of magnitude lower than in the EMEP scheme than in the Wesely scheme. From measurements of ozone fluxes in the Sahara desert, $v^\chem{O_3}_\text{desert, day} = 0.1\,\unit{cm\,s^{-1}}$, $v^\chem{O_3}_\text{desert, night} = 0.04\,\unit{cm\,s^{-1}}$, and $\left<v^\chem{O_3}_\text{desert}\right> = 0.065\,\unit{cm\,s^{-1}}$ have been deduced \citep{AE:Gusten1995}. This implies, that ozone dry deposition over desert regions is highly overestimated in the Wesely scheme, while it may be underestimated in the EMEP scheme. Similarly, the dry deposition velocities over water differ. From measurements over the ocean a value of $v^\chem{O_3}_\text{water} = 0.019\,\unit{cm\,s^{-1}}$ has been deduced \citep{JGR:Helmig2012}. In a comparison of various dry deposition mechanisms over ocean waters implemented in their model, \citet{ACP:Luhar2017} give a range of $0.018\,\unit{cm\,s^{-1}} \leq \left<v^\chem{O_3}_\text{water}\right> \leq 0.039\,\unit{cm\,s^{-1}}$.
%
\begin{figure*}[t]
  \includegraphics[width=12cm]{pictures/final-total_ozone_drydepvelo_monthly_pft}
  \caption{Average ozone dry deposition velocities separated by land use type and compared to results from \citep{ACP:Hardacre2015}.}
  \label{fig:mmm_drydep_season_pft}
\end{figure*}
%
